Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-02 paper.tex
More work on explaining flow behavior.

When mapping flow behaviors into a vector space, it is important that the mapping of behaviors into the vector spaces in question respect the vector space operations: addition of vectors and scalar multiplication. This requirement is best clarified with an intentionally simplistic example. Suppose that the behavior of a flow is represented by the number of packets, $p$, the total bytes transmitted, $b$, and the total duration, $d$. This triplet can be naturally embedded into Euclidean space: $\vec{x}=(p,b,d)\in\R^3$. If we have two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, then their ``sum'' behavior, $\vec{x}_1+\vec{x}_2$ has $p_1+p_2$ packets, $b_1+b_2$ bytes, $d_1+d_2$ duration. Their average, i.e. $\frac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.

@@

The {packets-bytes-duration} model representation is clearly quite simplistic, not to mention incomplete. It captures nothing about variability in packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; and it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: if the end-points, start time and end time of each flow is known, it suffices to know the empirical distribution of packet sizes and inter-packet intervals in order to reproduce network performance characteristics accurately. Accordingly, we represent a flow's behavior by these precise characteristics. We encode a single flow's behavior with six component vectors:
\begin{itemize}
\item $\vec{z} \in \R^{d_z}$ is a histogram of packet sizes;
\item $\vec{v} \in \R^{d_v}$ is a histogram of inter-packet intervals;
\item $\vec{b} \in \R^{d_t}$ indicates when the flow begins;
\item $\vec{e} \in \R^{d_t}$ indicates when the flow ends;
\item $\vec{s}\in\R^n$ indicates the source node of the flow;
\item $\vec{d}\in\R^n$ indicates the destination node of the flow.
\end{itemize}
The behavior vectors for all flows are collected into six corresponding matrices, whose rows are the behavior vectors above:
\begin{itemize}
\item $Z\in\R^{f \times d_z}$ the size matrix;
\item $V\in\R^{f \times d_v}$ the interval matrix;
\item $B\in\R^{f \times d_t}$ the start time matrix;
\item $E\in\R^{f \times d_t}$ the end time matrix;
\item $S\in\R^{f \times n}$ the source matrix;
\item $D\in\R^{f \times n}$ the destination matrix.
\end{itemize}

>>>>

When mapping flow behaviors into a vector space, it is important that the mapping of behaviors into the vector spaces in question respect the vector space operations: addition of vectors and scalar multiplication. This requirement is best clarified with a simple example. Suppose that the behavior of a flow is represented by the number of packets, $p$, the total bytes transmitted, $b$, and its total duration, $d$. This triplet can be naturally embedded into Euclidean space: $\vec{x}=(p,b,d)\in\R^3$. If we have two flows with behavior vectors $\vec{x}_1$ and $\vec{x}_2$, then their ``sum'' behavior, $\vec{x}_1+\vec{x}_2$ has $p_1+p_2$ packets, $b_1+b_2$ bytes, $d_1+d_2$ duration. Their average, i.e. $\frac{1}{2}(\vec{x}_1+\vec{x}_2)$, has the average number of packets, bytes, and duration, which is precisely what the ``average behavior'' of the two flows should intuitively be.

@@

The {packets-bytes-duration} model representation is clearly quite simplistic, not to mention incomplete. It captures nothing about variability in packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; and it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to aggregate behaviors of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: if the end-points, start time and end time of each flow are known, it suffices to know empirical distributions of packet sizes and inter-packet intervals for each flow to reproduce network performance accurately. Accordingly, we represent a flow's behavior by these precise characteristics. We encode a single flow's behavior with six component vectors:
\begin{itemize}
\item $\vec{z} \in \R^{d_z}$ is a histogram of packet sizes;
\item $\vec{v} \in \R^{d_v}$ is a histogram of inter-packet intervals;
\item $\vec{b} \in \R^{d_t}$ indicates when the flow begins;
\item $\vec{e} \in \R^{d_t}$ indicates when the flow ends;
\item $\vec{s}\in\R^n$ indicates the source node of the flow;
\item $\vec{d}\in\R^n$ indicates the destination node of the flow.
\end{itemize}
The behavior vectors for all flows are collected into six corresponding matrices, whose rows are the behavior vectors above:
\begin{itemize}
\item $Z\in\R^{f \times d_z}$ the size matrix;
\item $V\in\R^{f \times d_v}$ the interval matrix;
\item $B\in\R^{f \times d_t}$ the start time matrix;
\item $E\in\R^{f \times d_t}$ the end time matrix;
\item $S\in\R^{f \times n}$ the source matrix;
\item $D\in\R^{f \times n}$ the destination matrix.
\end{itemize}
